\documentclass{article}
\title{How to Speed Read}
\author{Grayson Wills}
\begin{document}
\maketitle


\section*{Introduction}
    \footnote{https://www.youtube.com/watch?v=e2-ahs905MQ}
    The speaker starts by saying that he has a reading speed of 1500 words per minute, 7X faster than what the average person reads. He claims that this is a skill, not a superpower, and that people can learn it relatively easily!
    
\section*{Issues Preventing Speed Reading}
    \begin{enumerate}
    \item Human beings stop being taught how to read at 12 years old, yet information comes at us faster and faster and also with more gravity behind it. He refers to this as \textbf{Information Overload}

    \item He introduces the concept of vocalization, where the reader thinks about what they are reading rather than seeing it. Humans can speak at about 250 words per minute, making the reading speed around 250 words per minute

    \end{enumerate}

\section*{Tips}
    \begin{enumerate}
    
        \item Press your tongue to the roof of your mouth. It stops you from mouthing the words, and it gives your brain something to think about

        \item Your environment matters a bunch. Put yourself in an environment where you can focus. Listen to classical music, or study in a place that is consistent and personable

        \item Don't go back when you're reading. Regression does nothing for yourself. It's your job to look for key figures

        \item Use an object to cover up the sentence you just read

        \item Don't fixate on the page

        \item Use a pacer to increase fixations

        \item Bouncing: meaning dividing your book into columns. Kind of like separating the sentence and reading between the spaces

        \item Read the front and back cover

        \item Read the Table of contents

        \item Skim: Look out for key subheadings and other things that standout

        \item Scan: Look around the items and diagrams you saw and look for key words

        \item Read it
    \end{enumerate}
    

\end{document}